---
title: "Softening"
author: "Tarrik Quneibi"
date: "7/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results=FALSE, fig.width=11, fig.height=7)

```

All packages needed to run the code.

``` {r libraries}

library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(ggforce)
library(tidyquant)
library(patchwork)
library(gridBase)
library(cowplot)
```

Importing the data and reading the fike to a variable.

```{r data import}
##setwd("`/R code/softening/")
softening_file <- list.files(pattern = "softening")
file <- read.csv(softening_file)
weighted_file <- list.files(pattern = "weighted")
raw_water <- read.csv(weighted_file)
color_list <- list.files(pattern = "color")
color <- read.csv(color_list)

```


This section cleans the data. Removes all NAs and empty cells. Converts all columns to their correct class. Combines the duplicate columns for hardness, calcium, and magnesium. Converts the calcium and magnesium columns to units of CaCO3. Removes all duplicate columns.

```{r Clean up}
file[is.na(file)] <- 0
file[file==""]<- 0

file[,4:36] <- sapply(file[,4:35],as.numeric)
file[is.na(file)] <- 0


file$Hardness <- file$X.DW_APC_SUB_Total.Hardness.by.2340B + file$HARDNESS
file$Calcium <- file$X.DW_APC_SUB_Calcium + file$CALCIUM+file$CHEM_SCADA_CA
file$Magnesium <- file$X.DW_APC_SUB_Magnesium + file$MAGNESIUM
file$Lime <- file$Total.Lime/file$Total.Flow
file$Calcium <- file$Calcium*2.48892
file$Magnesium <- file$Magnesium*4.11438

file$X.DW_APC_SUB_Total.Hardness.by.2340B <- NULL
file$HARDNESS <- NULL
file$X.DW_APC_SUB_Calcium <- NULL
file$CALCIUM <- NULL
file$CHEM_SCADA_CA <- NULL
file$X.DW_APC_SUB_Magnesium <- NULL
file$MAGNESIUM <- NULL
file$Total.Lime <- NULL

```
This chunk removes all extreme outliers from the pH and hardness data. anything aboce a pH of 14 and any hardness above 1000 mg/l CaCO3 are removed.

```{r cleaning pH and hardness columns}
i=0
file[is.na(file)] <- 0
file$Collection.Date.Time <- as.POSIXct(file$Collection.Date.Time, format="%m/%d/%Y %H:%M")
for (row in 1:nrow(file)) {
    Mag <- file[row, "Magnesium"]
    Date  <- file[row, "Collection.Date.Time"]

    if(Date > 2021-10-21) {
       file[row, "Magnesium"] <- (file[row, "Magnesium"])/2.03204
    }
}

for (PH in file$PH){
  i=i+1
  if (PH > 14) {
    file$PH[i] <- NA
  ##file$PH[i] <- 0
  }
}

i=0
for (Hardness in file$Hardness){
  i=i+1
  if (Hardness > 1000){
    file$Hardness[i] <- NA
}
}
```

```{r getting weighted raw data}
scada_data <- subset(file, subset = Location.description %in% c("Drinking Water SCADA Data"))
raw_river <- subset(file, subset = Location.description %in% c("WTP Raw River"))
raw_well <- subset(file, subset = Location.description %in% c("WTP Raw Well  Combined"))



Flow_data <- cbind.data.frame(scada_data$Collection.Date.Time, scada_data$FLOW_SCADA_PLANT1_RIVER)
Flow_data <- cbind.data.frame(Flow_data, scada_data$FLOW_SCADA_PLANT1_WELL)
Flow_data <- cbind.data.frame(Flow_data, scada_data$FLOW_SCADA_PLANT2_RIVER)
Flow_data <- cbind.data.frame(Flow_data, scada_data$FLOW_SCADA_PLANT2_WELL)
Flow_data$Total.Flow <- Flow_data$FLOW_SCADA_PLANT1_RIVER+ Flow_data$FLOW_SCADA_PLANT1_WELL + Flow_data$FLOW_SCADA_PLANT2_RIVER + Flow_data$FLOW_SCADA_PLANT2_WELL 
```
Organizes the full data set into seperate dataframes based on the loccation. Converts all date columns to date time class.

```{r organizing location bins}
basin2 <- data.frame()
basin3 <- data.frame()
basin4 <- data.frame()
basin5 <- data.frame()
reservoir <- data.frame()

basin2 <- subset(file, subset = Location.description %in% c("WTP Plant1 Basin2"))
basin3 <- subset(file, subset = Location.description %in% c("WTP Plant1 Basin3" ))
basin4 <- subset(file, subset = Location.description %in% c("WTP Plant2 Basin4"))
basin5 <- subset(file, subset = Location.description %in% c("WTP Plant2 Basin5"))
reservoir <- subset(file, subset = Location.description %in% c("WTP Main Reservoir"))

basin2$Location.description <- "Basin2"
basin3$Location.description <- "Basin3"
basin4$Location.description <- "Basin4"
basin5$Location.description <- "Basin5"
reservoir$Location.description <- "Reservoir"
basin2$Collection.Date.Time <- as.POSIXct(basin2$Collection.Date.Time,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
basin3$Collection.Date.Time <- as.POSIXct(basin3$Collection.Date.Time,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
basin4$Collection.Date.Time <- as.POSIXct(basin4$Collection.Date.Time,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
basin5$Collection.Date.Time <- as.POSIXct(basin5$Collection.Date.Time,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
reservoir$Collection.Date.Time <- as.POSIXct(reservoir$Collection.Date.Time,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
```

```{r subset for alk under 190}
file.alk.b190 <- subset(file, ALK_TOTAL < 190)
river_alk <- subset(file.alk.b190, subset = Location.description %in% c("WTP Raw River"))
color_river <- subset(color, subset = Location.description %in% c("WTP Raw River"))

color_river$COLOR <- as.numeric(color_river$COLOR)
color_river <- cbind.data.frame(color_river$Collection.Date.Time, color_river$COLOR)
colnames(color_river) <- c("Collection.Date.Time", "COLOR")
river_alk$Collection.Date.Time <- as.POSIXct(river_alk$Collection.Date.Time,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
color_river$Collection.Date.Time <- as.POSIXct(color_river$Collection.Date.Time,format="%m/%d/%Y %H:%M",tz=Sys.timezone())

alk_color <- left_join(river_alk, color_river, by = "Collection.Date.Time")

ggplot() + 
    geom_point(data = alk_color, aes(x=COLOR, y = ALK_TOTAL, color='River Alk')) +
    scale_colour_manual("", 
                      breaks = c("River Alk"),
                      values = c("blue")) +
   ggtitle("Alkalinity against river color")

ggplot() + 
    geom_point(data = river_alk, aes(x=Collection.Date.Time, y = ALK_TOTAL, color='River Alk')) +
    geom_point(data = color_river, aes(x=Collection.Date.Time, y = COLOR, color='River color')) +
    scale_colour_manual("", 
                      breaks = c("River Alk", "River color"),
                      values = c("blue", "red")) +
   ggtitle("Alkalinity against river color over time")+
  ylim(5,190)

```
```{r subset for color above 45}

```

This plots the hardness of all basins over time.

```{r scatter plotting hardness}
ggplot() + 
    geom_point(data = basin2, aes(x=PH, y = Hardness, color='Basin 2')) +
    geom_point(data = basin3, aes(x=PH, y = Hardness, color='Basin 3')) +
    geom_point(data = basin4, aes(x=PH, y = Hardness, color='Basin 4')) +
    geom_point(data = basin5, aes(x=PH, y = Hardness, color='Basin 5')) +
    scale_colour_manual("", 
                      breaks = c("Basin 2", "Basin 3", "Basin 4", "Basin 5"),
                      values = c("red", "green", "purple", "blue")) +
   ggtitle("Hardness against pH")+
   xlim(7.5 , 12.5)+
   ylim(50, 300)
#############################################################################################
Basin_4 <- cbind.data.frame(as.POSIXct(basin4$Collection.Date.Time, format = "%m/%d/%Y %H:%M"), basin4$PH)
colnames(Basin_4) <- c("Date", "pH_B4")
Basin_5 <- cbind.data.frame(as.POSIXct(basin5$Collection.Date.Time, format = "%m/%d/%Y %H:%M"), basin5$Hardness)
colnames(Basin_5) <- c("Date", "Hardness_B5")
Raw <- cbind.data.frame(as.POSIXct(raw_water$Collection.Date.Time,format = "%m/%d/%Y %H:%M"), raw_water$Hardness)
colnames(Raw) <- c("Date", "Hardness_R")

data1 <- left_join(Basin_5, Raw, by = "Date")
data1 <- left_join(data1, Basin_4, by= "Date")
data1[data1 == 0] <- NA
data1 <- data1[complete.cases(data1), ]

data1$Difference <- data1$Hardness_R - data1$Hardness_B5
#############################################################################################
 plot1 <- ggplot() + 
  geom_point(data = data1, aes(x = Date, y = Difference, color = "Raw - Basin 5")) +
  scale_colour_manual("", 
                      breaks = c("Raw - Basin 5"),
                      values = c("red")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ggtitle('Differences of hardness')+
  ylim(100, 300)


plot2 <- ggplot() + 
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = PH, color = "Basin 4")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4"),
                      values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') +
  ylab('pH')+
  ylim(9,12.5)+
  ggtitle('pH over time')+
  scale_x_datetime(date_breaks = "3 months")

plot1/plot2
################################################################################################
ggplot() + 
    geom_point(data = data1, aes(x=pH_B4, y = Difference, color='Raw - Basin 5')) +
    scale_colour_manual("", 
                      breaks = c("Raw - Basin 5"),
                      values = c("blue")) +
   ggtitle("Difference of Hardness against pH")+
   xlab("Basin 4 pH")+
   ylab("Hardness")

```

This plots all locations magnesium and calium concentrations against the pH. The last section plots the magnesium removed by basins 2 and 4 against their respective pH.

```{r scatter plot Ca and Mg against pH}

plot3 <- ggplot() + 
    geom_point(data = basin2, aes(x=PH, y = Calcium, color='Basin 2')) +
    geom_point(data = basin3, aes(x=PH, y = Calcium, color='Basin 3')) +
    geom_point(data = basin4, aes(x=PH, y = Calcium, color='Basin 4')) +
    geom_point(data = basin5, aes(x=PH, y = Calcium, color='Basin 5')) +
    geom_point(data = reservoir, aes(x=PH, y = Calcium, color='Reservoir')) +
    geom_point(data = raw_water, aes(x=PH, y = Calcium, color='Raw')) +
    scale_colour_manual("", 
                      breaks = c("Basin 2", "Basin 3", "Basin 4", "Basin 5", "Raw", "Reservoir"),
                      values = c("red", "green", "blue", "purple", "orange", "yellow")) +
   ggtitle("Calcium against pH")+
   xlim(7.5, 12.5)+
   ylim(5, 400)

plot4 <- ggplot() + 
    geom_point(data = basin2, aes(x=PH, y = Magnesium, color='Basin 2'), alpha=0.5) +
    geom_point(data = basin3, aes(x=PH, y = Magnesium, color='Basin 3'), alpha=0.5) +
    geom_point(data = basin4, aes(x=PH, y = Magnesium, color='Basin 4'), alpha=0.5) +
    geom_point(data = basin5, aes(x=PH, y = Magnesium, color='Basin 5'), alpha=0.5) +
    geom_point(data = reservoir, aes(x=PH, y = Magnesium, color='Reservoir')) +
    geom_point(data = raw_water, aes(x=PH, y = Magnesium, color='Raw')) +
    scale_colour_manual("", 
                      breaks = c("Basin 2", "Basin 3", "Basin 4", "Basin 5", "Raw", "Reservoir"),
                      values = c("red", "green", "blue", "purple", "orange", "yellow")) +
   ggtitle("Magnesium against pH")+
   xlim(6, 12.5)+
   ylim(5,300)
################################################################################################
Basin_2 <- cbind.data.frame(as.POSIXct(basin2$Collection.Date.Time, format = "%m/%d/%Y %H:%M") , basin2$Magnesium)
Basin_2 <- cbind.data.frame(Basin_2 , basin2$PH)
colnames(Basin_2) <- c("Date", "Magnesium_B2", "pH_B2")

Basin_4 <- cbind.data.frame(as.POSIXct(basin4$Collection.Date.Time, format = "%m/%d/%Y %H:%M"), basin4$Magnesium)
Basin_4 <- cbind.data.frame(Basin_4 , basin4$PH)
colnames(Basin_4) <- c("Date", "Magnesium_B4","pH_B4")

Raw <- cbind.data.frame(as.POSIXct(raw_water$Collection.Date.Time,format = "%m/%d/%Y %H:%M"), raw_water$Magnesium)
Raw <- cbind.data.frame(Raw , raw_water$PH)
colnames(Raw) <- c("Date", "Magnesium_R", "pH_R")

data1 <- left_join(Basin_2, Raw, by = "Date")
data2 <- left_join(Basin_4, Raw, by = "Date")
data1[data1 == 0] <- NA
data2[data2 == 0] <- NA

data1 <- data1[complete.cases(data1), ]
data2 <- data2[complete.cases(data2), ]

data1$Difference <- data1$Magnesium_R - data1$Magnesium_B2
data2$Difference <- data2$Magnesium_R - data2$Magnesium_B4
################################################################################################
plot5 <- ggplot() + 
  geom_point(data = data1, aes(x = pH_B2, y = Difference, color = "Raw - Basin 2")) +
  geom_point(data = data2, aes(x = pH_B4, y = Difference, color = "Raw - Basin 4")) +
  scale_colour_manual("", 
                      breaks = c("Raw - Basin 2", "Raw - Basin 4"),
                      values = c("red", "green"))+
  xlab('pH') +
  ylab('Magnesium (mg/l CaCO3')+
  ggtitle('Differences of Magnesium')
################################################################################################

```

This plots a boxplot of the calcium and magnesium concentrations separated by location.

```{r boxplot of Ca and Mg}
all_data <- bind_rows(basin2, basin3)
all_data <- bind_rows(all_data, basin4)
all_data <- bind_rows(all_data, basin5)
all_data <- bind_rows(all_data, raw_water)
data <- data.frame()
data <- cbind(all_data$Location.description, all_data$Collection.Date.Time)
data <- cbind(data, all_data$Calcium)
data <- cbind(data, all_data$Magnesium)
data <- as.data.frame(data)
colnames(data) <- c("Location", "Date", "Calcium", "Magnesium")
data$Location <- as.factor(data$Location)
data$Calcium <- as.numeric(data$Calcium)
data$Magnesium <- as.numeric(data$Magnesium)
data$Date <- as.POSIXct(data$Date,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
data[data == 0] <- NA

boxplot(data$Calcium~data$Location, col=rainbow(4),
        main = "Calcium concentration at each location in the treatment train",
           xlab = "Location",
           ylab = "Calcium (mg/l CaCO3)")
boxplot(data$Magnesium~data$Location, col=rainbow(4),
        main = "Magnesium concentration at each location in the treatment train",
           xlab = "Location",
           ylab = "Magnesium (mg/l CaCO3)")
```

This plots a boxplot of the hardness at each location.

```{r box plot of hardness}
all_data <- bind_rows(basin2, basin3)
all_data <- bind_rows(all_data, basin4)
all_data <- bind_rows(all_data, basin5)
all_data <- bind_rows(all_data, raw_water)
data <- data.frame()
data <- cbind(all_data$Location.description, all_data$Collection.Date.Time)
data <- cbind(data, all_data$Hardness)
data <- as.data.frame(data)
colnames(data) <- c("Location", "Date", "Hardness")
data$Location <- as.factor(data$Location)
data$Hardness <- as.numeric(data$Hardness)
data$Date <- as.POSIXct(data$Date,format="%m/%d/%Y %H:%M",tz=Sys.timezone())
data[data == 0] <- NA

boxplot(data$Hardness~data$Location, col=rainbow(4),
        main = "Total hardness at each location in the treatment train",
           xlab = "Location",
           ylab = "Hardness (mg/l CaCO3)")

```

This plots all locations hardness over time.

```{r hardness over time}

ggplot() + 
  geom_point(data = basin2, aes(x=Collection.Date.Time, y = Hardness, color='Basin 2')) +
    geom_point(data = basin3, aes(x=Collection.Date.Time, y = Hardness, color='Basin 3')) +
    geom_point(data = basin4, aes(x=Collection.Date.Time, y = Hardness, color='Basin 4')) +
    geom_point(data = basin5, aes(x=Collection.Date.Time, y = Hardness, color='Basin 5')) +
    geom_point(data = raw_water, aes(x=Collection.Date.Time, y = Hardness, color='Raw')) +
    scale_colour_manual("", 
                      breaks = c("Basin 2", "Basin 3", "Basin 4", "Basin 5", "Raw"),
                      values = c("red", "green", "blue", "purple", "orange")) +
  ylim(50, 450)+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('Hardness (mg/l CaCO3)') +
  ggtitle('Hardness over time for all basins')



```

This plots the hardness removed by each basin over time. Also plots for each individual year.

```{r difference of primary and secondary hardness}
Basin_2 <- cbind.data.frame(as.POSIXct(basin2$Collection.Date.Time, format = "%m/%d/%Y %H:%M") , basin2$Hardness)
colnames(Basin_2) <- c("Date", "Hardness_B2")
Basin_3 <- cbind.data.frame(as.POSIXct(basin3$Collection.Date.Time, format = "%m/%d/%Y %H:%M"), basin3$Hardness)
colnames(Basin_3) <- c("Date", "Hardness_B3")
Basin_4 <- cbind.data.frame(as.POSIXct(basin4$Collection.Date.Time, format = "%m/%d/%Y %H:%M"), basin4$Hardness)
colnames(Basin_4) <- c("Date", "Hardness_B4")
Basin_5 <- cbind.data.frame(as.POSIXct(basin5$Collection.Date.Time, format = "%m/%d/%Y %H:%M"), basin5$Hardness)
colnames(Basin_5) <- c("Date", "Hardness_B5")
Raw <- cbind.data.frame(as.POSIXct(raw_water$Collection.Date.Time,format = "%m/%d/%Y %H:%M"), raw_water$Hardness)
colnames(Raw) <- c("Date", "Hardness_R")

data1 <- left_join(Basin_2, Basin_3, by = "Date")
data2 <- left_join(Basin_4, Basin_5, by = "Date")
data3 <- left_join(Basin_2, Raw, by = "Date")
data4 <- left_join(Basin_3, Raw, by = "Date")
data1[data1 == 0] <- NA
data2[data2 == 0] <- NA
data3[data3 == 0] <- NA
data4[data4 == 0] <- NA
data1 <- data1[complete.cases(data1), ]
data2 <- data2[complete.cases(data2), ]
data3 <- data3[complete.cases(data3), ]
data4 <- data4[complete.cases(data4), ]

data1$Difference <- data1$Hardness_B2 - data1$Hardness_B3
data2$Difference <- data2$Hardness_B4 - data2$Hardness_B5
data3$Difference <- data3$Hardness_R - data3$Hardness_B2
data4$Difference <- data4$Hardness_R - data4$Hardness_B3

####################################################################################################
ggplot() + 
  geom_point(data = data1, aes(x = Date, y = Difference, color = "Basin 2 - Basin 3")) +
  geom_point(data = data2, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5"), alpha = 0.2) +
  geom_point(data = data3, aes(x = Date, y = Difference, color = "Raw - Basin 2")) +
  geom_point(data = data4, aes(x = Date, y = Difference, color = "Raw - Basin 3")) +
  scale_colour_manual("", 
                      breaks = c("Basin 2 - Basin 3", "Basin 4 - Basin 5" ,"Raw - Basin 2", "Raw - Basin 3" ),
                      values = c("red", "green", "blue", "purple")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('Hardness (mg/l CaCO3')+
  ggtitle('Differences of hardness')
####################################################################################################
ggplot() + 
  geom_point(data = data1, aes(x = Date, y = Difference, color = "Basin 2 - Basin 3")) +
  geom_point(data = data2, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5"), alpha = 0.2) +
  geom_point(data = data3, aes(x = Date, y = Difference, color = "Raw - Basin 2")) +
  geom_point(data = data4, aes(x = Date, y = Difference, color = "Raw - Basin 3")) +
  scale_colour_manual("", 
                      breaks = c("Basin 2 - Basin 3", "Basin 4 - Basin 5" ,"Raw - Basin 2", "Raw - Basin 3" ),
                      values = c("red", "green", "blue", "purple")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "1 month")+
  xlab('Date') +
  ylab('Hardness (mg/l CaCO3')+
  ggtitle('Differences of hardness in 2018')+
  xlim(as.POSIXct('2018-01-01'),as.POSIXct('2019-01-01'))
#####################################################################################################
ggplot() + 
  geom_point(data = data1, aes(x = Date, y = Difference, color = "Basin 2 - Basin 3")) +
  geom_point(data = data2, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5"), alpha = 0.2) +
  geom_point(data = data3, aes(x = Date, y = Difference, color = "Raw - Basin 2")) +
  geom_point(data = data4, aes(x = Date, y = Difference, color = "Raw - Basin 3")) +
  scale_colour_manual("", 
                      breaks = c("Basin 2 - Basin 3", "Basin 4 - Basin 5" ,"Raw - Basin 2", "Raw - Basin 3" ),
                      values = c("red", "green", "blue", "purple")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "1 month")+
  xlab('Date') +
  ylab('Hardness (mg/l CaCO3')+
  ggtitle('Differences of hardness in 2019')+
  xlim(as.POSIXct('2019-01-01'),as.POSIXct('2020-01-01'))
#####################################################################################################
ggplot() + 
  geom_point(data = data1, aes(x = Date, y = Difference, color = "Basin 2 - Basin 3")) +
  geom_point(data = data2, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5"), alpha = 0.2) +
  geom_point(data = data3, aes(x = Date, y = Difference, color = "Raw - Basin 2")) +
  geom_point(data = data4, aes(x = Date, y = Difference, color = "Raw - Basin 3")) +
  scale_colour_manual("", 
                      breaks = c("Basin 2 - Basin 3", "Basin 4 - Basin 5" ,"Raw - Basin 2", "Raw - Basin 3" ),
                      values = c("red", "green", "blue", "purple")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "1 month")+
  xlab('Date') +
  ylab('Hardness (mg/l CaCO3')+
  ggtitle('Differences of hardness in 2020')+
  xlim(as.POSIXct('2020-01-01'),as.POSIXct('2021-01-01'))
#####################################################################################################
ggplot() + 
  geom_point(data = data1, aes(x = Date, y = Difference, color = "Basin 2 - Basin 3")) +
  geom_point(data = data2, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5"), alpha = 0.2) +
  geom_point(data = data3, aes(x = Date, y = Difference, color = "Raw - Basin 2")) +
  geom_point(data = data4, aes(x = Date, y = Difference, color = "Raw - Basin 3")) +
  scale_colour_manual("", 
                      breaks = c("Basin 2 - Basin 3", "Basin 4 - Basin 5" ,"Raw - Basin 2", "Raw - Basin 3" ),
                      values = c("red", "green", "blue", "purple")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "1 month")+
  xlab('Date') +
  ylab('Hardness (mg/l CaCO3')+
  ggtitle('Differences of hardness in 2021')+
  xlim(as.POSIXct('2021-01-01'),as.POSIXct('2021-07-01'))
#####################################################################################################

```

This plots the difference in hardness between basin 5 and 4, the alkalinity of both basins, the pH of both basins, and the temperature of the reservoir. Also plots for each individual year.


```{r differences by year}


data2.bel10 <- subset(data2, Difference < 10)
reservoir$TEMP_C <- as.numeric(reservoir$TEMP_C)
reservoir_T <- cbind.data.frame(as.POSIXct(reservoir$Collection.Date.Time, format = "%m/%d/%Y %H:%M") , reservoir$TEMP_C)
reservoir_T[reservoir_T == 0] <- NA
basin4[basin4 == 0] <- NA
basin5[basin5 == 0] <- NA
colnames(reservoir_T) <- c("Date", "Temp")
##################################################################################################### All data
plot1 <- ggplot() + 
  geom_point(data = data2.bel10, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +

  geom_point(data = reservoir_T, aes(x = Date, y = Temp, color = "Reservoir temperature")) +
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 4 alkalinity")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 5 alkalinity")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4 - Basin 5", "Reservoir temperature", "Basin 4 alkalinity", "Basin 5 alkalinity" ),
                      values = c("red", "purple", "orange", "yellow")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylim(-50, 175)+
  ggtitle('Differences of hardness against pH, alkalinity, and temperature')

plot2 <- ggplot() + 
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = PH, color = "Basin 4")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = PH, color = "Basin 5")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4", "Basin 5"),
                      values = c("red","blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('pH')+
  ylim(9,12.5)+
  ggtitle('pH over time')

plot1/plot2
##################################################################################################### For 2018
plot1 <- ggplot() + 
  geom_point(data = data2.bel10, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +

  geom_point(data = reservoir_T, aes(x = Date, y = Temp, color = "Reservoir temperature")) +
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 4 alkalinity")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 5 alkalinity")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4 - Basin 5", "Reservoir temperature", "Basin 4 alkalinity", "Basin 5 alkalinity" ),
                      values = c("red", "purple", "orange", "yellow")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  ##scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ggtitle('Differences of hardness against pH, alkalinity, and temperature in 2018')+
  ylim(-50, 175)+
  xlim(as.POSIXct('2018-01-01'),as.POSIXct('2019-01-01'))

plot2 <- ggplot() + 
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = PH, color = "Basin 4")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = PH, color = "Basin 5")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4", "Basin 5"),
                      values = c("red","blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  ##scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('pH')+
  ylim(9,12.5)+
  ggtitle('pH over time')+
  xlim(as.POSIXct('2018-01-01'),as.POSIXct('2019-01-01'))

plot1/plot2
##################################################################################################### For 2019
plot1 <- ggplot() + 
  geom_point(data = data2.bel10, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +

  geom_point(data = reservoir_T, aes(x = Date, y = Temp, color = "Reservoir temperature")) +
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 4 alkalinity")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 5 alkalinity")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4 - Basin 5", "Reservoir temperature", "Basin 4 alkalinity", "Basin 5 alkalinity" ),
                      values = c("red", "purple", "orange", "yellow")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  ##scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ggtitle('Differences of hardness against pH, alkalinity, and temperature in 2019')+
  ylim(-50, 175)+
  xlim(as.POSIXct('2019-01-01'),as.POSIXct('2020-01-01'))

plot2 <- ggplot() + 
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = PH, color = "Basin 4")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = PH, color = "Basin 5")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4", "Basin 5"),
                      values = c("red","blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  ##scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('pH')+
  ylim(9,12.5)+
  ggtitle('pH over time')+
  xlim(as.POSIXct('2019-01-01'),as.POSIXct('2020-01-01'))

plot1/plot2
##################################################################################################### For 2020
plot1 <- ggplot() + 
  geom_point(data = data2.bel10, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +

  geom_point(data = reservoir_T, aes(x = Date, y = Temp, color = "Reservoir temperature")) +
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 4 alkalinity")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 5 alkalinity")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4 - Basin 5", "Reservoir temperature", "Basin 4 alkalinity", "Basin 5 alkalinity" ),
                      values = c("red", "purple", "orange", "yellow")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
 ## scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ggtitle('Differences of hardness against pH, alkalinity, and temperature in 2020')+
  ylim(-50, 175)+
  xlim(as.POSIXct('2020-01-01'),as.POSIXct('2021-01-01'))

plot2 <- ggplot() + 
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = PH, color = "Basin 4")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = PH, color = "Basin 5")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4", "Basin 5"),
                      values = c("red","blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  ##scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('pH')+
  ylim(9,12.5)+
  ggtitle('pH over time')+
  xlim(as.POSIXct('2020-01-01'),as.POSIXct('2021-01-01'))

plot1/plot2
#####################################################################################################for 2021
plot1 <- ggplot() + 
  geom_point(data = data2.bel10, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  geom_point(data = reservoir_T, aes(x = Date, y = Temp, color = "Reservoir temperature")) +
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 4 alkalinity")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 5 alkalinity")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4 - Basin 5", "Reservoir temperature", "Basin 4 alkalinity", "Basin 5 alkalinity" ),
                      values = c("red", "purple", "orange", "yellow")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  ##scale_y_continuous(limits = c(-50,175), name = "mg/l CaCO3", sec.axis = sec_axis(~./14.29, name = "pH")) +
  ##scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylim(-50, 175)+
  ggtitle('Differences of hardness against pH, alkalinity, and temperature in 2021')+
  xlim(as.POSIXct('2021-01-01'),as.POSIXct('2021-07-01'))

plot2 <- ggplot() + 
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = PH, color = "Basin 4")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = PH, color = "Basin 5")) +
  scale_colour_manual("", 
                      breaks = c("Basin 4", "Basin 5"),
                      values = c("red","blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  ##scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('pH')+
  ylim(9,12.5)+
  ggtitle('pH over time')+
  xlim(as.POSIXct('2021-01-01'),as.POSIXct('2021-07-01'))

plot1/plot2
###################################################################################

Jan <- subset(data2, format.Date(Date, "%m")=="01")
Jan$Date <- "Jan"
Feb <- subset(data2, format.Date(Date, "%m")=="02")
Feb$Date <- "Feb"
Mar <- subset(data2, format.Date(Date, "%m")=="03")
Mar$Date <- "Mar"
Apr <- subset(data2, format.Date(Date, "%m")=="04")
Apr$Date <- "Apr"
May <- subset(data2, format.Date(Date, "%m")=="05")
May$Date <- "May"
Jun <- subset(data2, format.Date(Date, "%m")=="06")
Jun$Date <- "Jun"
Jul <- subset(data2, format.Date(Date, "%m")=="07")
Jul$Date <- "Jul"
Aug <- subset(data2, format.Date(Date, "%m")=="08")
Aug$Date <- "Aug"
Sep <- subset(data2, format.Date(Date, "%m")=="09")
Sep$Date <- "Sep"
Oct <- subset(data2, format.Date(Date, "%m")=="10")
Oct$Date <- "Oct"
Nov <- subset(data2, format.Date(Date, "%m")=="11")
Nov$Date <- "Nov"
Dec <- subset(data2, format.Date(Date, "%m")=="12")
Dec$Date <- "Dec"

month.list <- list(Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec)
month.data<- bind_rows(month.list, .id = "column_label")
month.data$Date <- factor(month.data$Date, c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
boxplot(month.data$Difference~month.data$Date, col=rainbow(4),
        main = "Difference in total hardness from basin 5 - basin 4",
           xlab = "Month",
           ylab = "Hardness (mg/l CaCO3)")


Jan <- subset(data2, format.Date(Date, "%m")=="01")
Feb <- subset(data2, format.Date(Date, "%m")=="02")
Mar <- subset(data2, format.Date(Date, "%m")=="03")
Apr <- subset(data2, format.Date(Date, "%m")=="04")
May <- subset(data2, format.Date(Date, "%m")=="05")
Jun <- subset(data2, format.Date(Date, "%m")=="06")
Jul <- subset(data2, format.Date(Date, "%m")=="07")
Aug <- subset(data2, format.Date(Date, "%m")=="08")
Sep <- subset(data2, format.Date(Date, "%m")=="09")
Oct <- subset(data2, format.Date(Date, "%m")=="10")
Nov <- subset(data2, format.Date(Date, "%m")=="11")
Dec <- subset(data2, format.Date(Date, "%m")=="12")

plot1 <- ggplot() + 
  geom_point(data = Jan, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in Janurary')

plot2 <- ggplot() + 
  geom_point(data = Feb, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in February')

plot3 <- ggplot() + 
  geom_point(data = Mar, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in March')

plot4 <- ggplot() + 
  geom_point(data = Apr, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in April')

plot5 <- ggplot() + 
  geom_point(data = May, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in May')

plot6 <- ggplot() + 
  geom_point(data = Jun, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in June')

plot7 <- ggplot() + 
  geom_point(data = Jul, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in July')

plot8 <- ggplot() + 
  geom_point(data = Aug, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in August')

plot9 <- ggplot() + 
  geom_point(data = Sep, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in September')

plot10 <- ggplot() + 
  geom_point(data = Oct, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in October')

plot11 <- ggplot() + 
  geom_point(data = Nov, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in November')

plot12 <- ggplot() + 
  geom_point(data = Dec, aes(x = Date, y = Difference, color = "Basin 4 - Basin 5")) +
  scale_colour_manual("", breaks = c("Basin 4 - Basin 5"), values = c("blue")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  xlab('Date') + ggtitle('Differences of hardness over time in December')
plot.list <- c(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, plot11, plot12)

(plot1+plot2+plot3)/(plot4+plot5+plot6)/(plot7+plot8+plot9)/(plot10+plot11+plot12)

```
Plots the alkalinity of all basins over time.

```{r Alkalinity over time}

ggplot() + 
  geom_point(data = basin2, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 2")) +
  geom_point(data = basin3, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 3")) +
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 4")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Basin 5")) +
  geom_point(data = reservoir, aes(x = Collection.Date.Time, y = ALK_TOTAL, color = "Reservoir")) +
  scale_colour_manual("", 
                      breaks = c("Basin 2", "Basin 3" ,"Basin 4", "Basin 5", "Reservoir" ),
                      values = c("red", "green", "blue", "purple", "orange")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('Alkalinity (mg/l CaCO3')+
  ylim(10, 350)+
  ggtitle('Alkalinity over time')


```

Plots the pH of allocations over time.

```{r pH over time}
ggplot() + 
  geom_point(data = basin2, aes(x = Collection.Date.Time, y = PH, color = "Basin 2")) +
  geom_point(data = basin3, aes(x = Collection.Date.Time, y = PH, color = "Basin 3")) +
  geom_point(data = basin4, aes(x = Collection.Date.Time, y = PH, color = "Basin 4")) +
  geom_point(data = basin5, aes(x = Collection.Date.Time, y = PH, color = "Basin 5")) +
  geom_point(data = reservoir, aes(x = Collection.Date.Time, y = PH, color = "Reservoir")) +
  geom_point(data = raw_water, aes(x = Collection.Date.Time, y = PH, color = "Raw")) +
  scale_colour_manual("", 
                      breaks = c("Basin 2", "Basin 3" ,"Basin 4", "Basin 5", "Reservoir", "Raw" ),
                      values = c("red", "green", "blue", "purple", "orange", "yellow")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  scale_x_datetime(date_breaks = "3 months")+
  xlab('Date') +
  ylab('pH')+
  ylim(5,12.5)+
  ggtitle('pH over time')
  

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
